Assessment of a Shen-Fourier turbulent channel flow solver for large-scale simulations

### 2016-01-29 paper.tex
Move to KMM solver

\section{Preliminaries}
The Navier-Stokes equations used to describe turbulent flow in a doubly periodic channel are
\begin{align}
 \frac{\partial \bm{u}}{\partial t} + (\bm{u} \cdot \nabla) \bm{u}   &= \nu \nabla^2 \bm{u} - \nabla{p} + \bm{f},  \label{eq:NS} \\
 \nabla \cdot \bm{u} &= 0,
\end{align}
where $\bm{u}(\bm{x}, t)=(u, v, w)$ is the velocity vector, $\bm{x}=(x, y, z)$ and $t$ are position and time,   $p(\bm{x}, t)$ is the instantaneous pressure normalized by a constant density and $\bm{f} = (0, f_y, 0)$ is a constant forcing vector. The computational domain is $\Omega=[-1, 1]\times [0, L_y] \times [0, L_z]$ with channel walls located at $x=\pm 1$ such that no-slip applies at $ \bm{u}(\pm 1, y, z, t) = 0$. The walls are spanning the $y-z$ plane and the equations are periodic in the $y$ and $z$ directions.

The Navier Stokes equations are discretized using Fourier basis functions for the periodic directions, and a combination of Chebyshev polynomials in the wall normal direction. Three different sets of basis functions and function spaces are relevant for the wall normal direction
\begin{align}
&  \phi_k(x) = T_k(x), & V_{N_x} &= \text{span}\{\phi_0, \phi_1,\ldots, \phi_{N_x}\} \label{eq:Tk}\\
& \D{\phi}_k(x) = T_k(x) - T_{k+2}(x), & \D{V}_{N_x} &= \text{span} \{ \D{\phi}_0, \D{\phi}_1, \ldots, \D{\phi}_{N_x-2} \} \label{eq:phiD}\\
& \N{\phi}_k(x) = T_k(x) - \left(\frac{k}{k+2}\right)^2 T_{k+2}(x), & \N{V}_{N_x} &= \text{span} \{\N{\phi}_1, \N{\phi}_2, \ldots, \N{\phi}_{N_x-2} \} \label{eq:phiN} 
\end{align}
where $T_k(x)$ is the $k$'th degree Chebyshev polynomial of the first kind. The basis functions and function spaces in (\ref{eq:phiD}) and (\ref{eq:phiN}) were suggested by Shen, and satisfy, respectively, the boundary conditions $\D{\phi}_k(\pm 1) = 0$ and $\N{\phi}'_k(\pm 1)=0$. The last two function spaces may alternatively be written as $\D{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1)=0 \}$ and $\N{V}_{N_x} = \{v \in V_{N_x}: \int_{-1}^{1}v/\sqrt{1-x^2} dx = 0, \, v'(\pm 1) = 0 \}$.

Three-dimensional basis functions that are periodic in $y$ and $z$ directions, can now be defined as
\begin{align}
  \psi_{\bm{k}}(\bm{x}) = \phi_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad V_N &= \text{span} \{ \psi_{\bm{k}}(\bm{x}):\, \bm{k} \in K_N  \}, \\
  \D{\psi}_{\bm{k}}(\bm{x}) = \D{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \D{V}_N &= \text{span} \{ \D{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \D{K}_N  \}, \\
  \N{\psi}_{\bm{k}}(\bm{x}) = \N{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \N{V}_N &= \text{span} \{ \N{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \N{K}_N  \},
\end{align}
where $\imath=\sqrt{-1}$ and 
\begin{align}
K_N = \Big\{\bm{k} \in \mathcal{Z}\times\mathcal{R}^2 / &(l, \underline{m}, \underline{n}) \in \left( l, \frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
 &(l, m, n) \in  [0, \ldots, N_x] \times [-\frac{N_y}{2},\ldots, \frac{N_y}{2}-1] \times[-\frac{N_z}{2},\ldots, \frac{N_z}{2}-1] \Big\}. \label{eq:wavenumbermesh}
\end{align}
$\D{K}_N$ and $\N{K}_N$ are both equal to $K_N$ except from in index $l$ that ranges from $[0, 1, \ldots, N_x-2]$ and $[1, 2, \ldots, N_x-2]$ for $\D{K}_N$ and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). The computational mesh has in real space been created using $N = (N_x, N_y, N_z)$ intervals, where the two periodic directions use uniform intervals. The computational mesh is given as
\begin{align}
  X_N = \Big\{ \bm{x} \in \mathcal{R}^3/&(x_i, y_j, z_k) \in \left( h(i), \frac{jL_y}{N_y}, \frac{kL_z}{N_z} \right), \text{where} \notag \\
  &(i, j, k) \in [0, 1, \ldots, N_x] \times [0, 1, \ldots, N_y-1] \times [0, 1, \ldots, N_z-1] \Big\} \label{eq:Xn}
\end{align}
where $x_i = h(i)$ represents either
\begin{equation}
 h(i) = \begin{cases}
   \cos \left(\frac{i \pi }{N_x} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Lobatto points}, \\
   \cos \left(\frac{(2i +1)\pi}{2N_x+2} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Chebyshev points}. \\
 \end{cases}
\end{equation}

For the Navier Stokes equations we look for solutions for the velocity components of the form
\begin{equation}
u(\bm{x}, t) = \frac{1}{N_yN_z}\sum_{\bm{k} \in \D{K}_N} \hat{u}_{\bm{k}}(t) \D{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_sol}
\end{equation}
where $\hat{u}_{\bm{k}}(t) = \hat{u}(l, {m}, {n}, t)$ are the expansion coefficients for the velocity component in $x$-direction and the scaling by $N_y$ and $N_z$ is merely for convenience and compliance with the inverse discrete Fourier transform. Similar expressions apply for the remaining two velocity components. Note that from now on we will simply use the notation $\hat{u}$ for $\hat{u}_{\bm{k}}(t)$, when it is possible to simplify without loss of clarity. Likewise we will simply use $u$ for $u(\bm{x}, t)$.

If the expansion coefficients $\hat{u}$  are known for the entire wavenumber mesh (\ref{eq:wavenumbermesh}), then the expression (\ref{eq:u_sol}) may be evaluated on the mesh (\ref{eq:Xn}) using fast Fourier and fast Shen inverse transforms as
\begin{align}
u(x_i, y_j, z_k, t) &= \underbrace{\frac{1}{N_z}\sum_n \underbrace{\frac{1}{N_y} \sum_m \underbrace{\sum_l \hat{u}(l,m,n)\D{\phi}_l(x_i)}_{\mathcal{S}_x^{-1}} e^{\imath \underline{m} y_j}}_{\mathcal{F}_y^{-1}} e^{\imath \underline{n} z_k}}_{\mathcal{F}_z^{-1}} , \quad \forall \, \bm{x} \in X_N, \notag \\
  u &= \D{\mathcal{T}}^{-1}(\hat{u}) =  \mathcal{F}^{-1}_z (\mathcal{F}^{-1}_y (\D{\mathcal{S}}^{-1}_x (\hat{u}))), \quad \forall \, \bm{x} \in X_N.  \label{eq:ifft} 
\end{align}
Here $\D{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse transform, $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represents inverse Fourier transforms along directions $y$ and $z$ respectively, and the inverse Shen transform $\D{\mathcal{S}}_{x}^{-1}$ is performed along the wall normal $x$ direction. Note that the transforms are applied sequentially to the entire computational mesh. The inverse Shen transform may be computed using fast Chebyshev transforms for all the three bases in (\ref{eq:Tk}, \ref{eq:phiD}, \ref{eq:phiN}). The transforms are slightly different for the three bases and ${\mathcal{S}}, \D{\mathcal{S}}$ and $\N{\mathcal{S}}$ are used to distinguish between them with obvious notation. Python implementations of forward and inverse Shen transforms and scalar products are given in the Appendix.

A three-dimensional scalar product in the weighted $[L^2_w(\Omega)]^3$ space is defined as
\begin{align}
 \left<u, v\right>_w &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}w\,dxdydz, \notag \\
   &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}\,dx_wdydz,
\end{align}
where $v^*$ is the comlex conjugate of $v$ and the weights $w$ are unity for periodic directions and  $w=1/\sqrt{1-x^2}$ for the inhomogeneous direction, and we have used $dx_w = dx/\sqrt{1-x^2}$. In this work we will make use of the discrete weighted $[l^2_w(\Omega)]^3$ space, where quadrature is employed for integration. As such the scalar product corresponds to sums that may be computed using fast transforms
\begin{align}
 \left< u, \D{\psi}_{\bm{k}} \right>_w &= h \underbrace{\sum_i \underbrace{\sum_j \underbrace{\sum_k u(x_i, y_j, z_k, t)  e^{-\imath \underline{n} z_k}}_{\mathcal{F}_n}  e^{-\imath \underline{m} y_j} }_{\mathcal{F}_m} \D{\phi}_{l}(x_i) w_i}_{\D{\mathcal{S}}_l},   \notag \\
  &=  h \D{\mathcal{S}}(u) = h \D{\mathcal{S}}_{l} (\mathcal{F}_{{m}} (\mathcal{F}_{{n}}(u))), \quad \forall \, \bm{k} \in \D{K}_N. \label{eq:sst1}
\end{align}
Here $h\D{\mathcal{S}}$ denotes the complete three dimensional scalar product and $h = L_yL_zN_y^{-1}N_z^{-1}$ is a constant. $\mathcal{F}_{{n}}$ and $\mathcal{F}_{{m}}$ represent discrete Fourier transforms in $z$- and $y$-directions, respectively, and $\D{\mathcal{S}}_{l}(\cdot) = (\cdot, \D{\phi}_l)_w$ is the Shen scalar product in the $x$-direction. The weights, $w_i$, required for the Shen transforms are defined as
\begin{equation}
 w_i = \begin{cases}
       \frac{\pi}{c_i N_x} &\forall \, i=0,1,\ldots, N_x \quad  \text{for Gauss-Lobatto},\\
       \frac{\pi}{N_x+1} &\forall \, i=0,1,\ldots, N_x  \quad \text{for Gauss-Chebyshev},      
 \end{cases}
\end{equation}
where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$.

The scalar product  $h\D{\mathcal{S}}$ in (\ref{eq:sst1}) does not represent a complete transform. To find a transformation between physical $u$ and spectral $\hat{u}$ we make use of Eq. (\ref{eq:u_sol}) directly and use the discrete orthogonality of the Fourier basis functions
\begin{align}
\left<u, \D{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< \sum_{\bm{m}=(q,r,s) \in \D{K}_N} \hat{u}_{\bm{m}} \D{\psi}_{\bm{m}}, \D{\psi}_{\bm{k}} \right>_w, \notag \\
           &= h \sum_{q=0}^{N_x-2} \sum_{i=0}^{N_x} \D{\phi}_{q}(x_i) \D{\phi}_{l}(x_i) w_i \, \hat{u}(q, {m}, {n}, t), \notag \\
           &= h \sum_{p=0}^{N_x-2} \D{B}_{lq} \hat{u}(q, {m}, {n}, t), \label{eq:sst2}
\end{align}
where the scalar product $\D{B}_{lq} = (\D{\phi}_q, \D{\phi}_l)_w = \sum_{i=0}^{N_x} \D{\phi}_{q}(x_i) \D{\phi}_{l}(x_i) w_i$ can be assembled to a tridiagonal mass matrix. The complete transformation is now obtained by setting Eq. (\ref{eq:sst2}) equal to (\ref{eq:sst1}) and solving for $\hat{u}$. Moving to matrix notation we get
\begin{align}
h\sum_{q=0}^{N_x-2} \D{B}_{lq} \hat{u}(q, {m}, {n}, t) &= h\D{\mathcal{S}}(u)(l, m, n, t), \notag \\
 \hat{u}_{\bm{k}}(t) &= \D{\mathcal{T}}(u) =  \D{B}^{-1}\D{\mathcal{S}}(u), \quad \forall \, \bm{k} \in \D{K}_N.
\end{align}
where $\D{\mathcal{T}}(u)$ denotes the complete transformation, such that $u = \D{\mathcal{T}}^{-1}(\D{\mathcal{T}}(u))$. Note that since $\D{B}$ assembles to a tridiagonal matrix the solution can be obtained very fast ($O(N_x)$) using one forward elimination followed by a backwards substitution (Thomas algorithm). The complete transformation thus requires a fast Chebyshev transform and a fast linear algebraic solve. Similar transforms $\mathcal{T}$ and $\N{\mathcal{T}}$ are defined for the two other bases (\ref{eq:Tk}) and (\ref{eq:phiN}), using mass matrices $B_{kj}=(T_j, T_k)_w$ and $\N{B}_{kj}=(\N{\phi}_j, \N{\phi}_j)_w$ and scalar products $\mathcal{S}_k(\cdot) = (\cdot, T_k)_w$ and $\N{\mathcal{S}}_k(\cdot) = (\cdot, \N{\phi}_k)_w$.

\section{The incremental pressure correction scheme}

The Navier Stokes equations are discretized in time using a second order incremental pressure correction scheme with Crank-Nicolson on diffusion and explicit Adams Bashforth on the nonlinear convection. The incremental pressure correction scheme (IPCS) that is used consists of a tentative momentum equation followed by a pressure correction. The two steps may be performed iteratively. The equations are iterated forward in time in equal size time steps $\triangle t$ from initial conditions $u(\bm{x}, 0) = u^0(\bm{x})$ at $t=0$, such that $t_n = n\triangle t$, $n \in \mathcal{Z}$. For turbulent channel flows initial conditions are simply a perturbed state used to get a turbulent flow going.

The IPCS attempts to find velocity vector $\bm{u}^{n}$ on time step $t_n$  and pressure midway between $t_n$ and $t_{n-1}$, $p^{n-1/2}$, given the solutions on all previous time steps. The three steps that are required solved on each time step are
\begin{align}
i)&\,\, \frac{\bm{u}^{*}-\bm{u}^{n-1}}{\triangle t} + \bm{N}^{n-1/2}   = \nu \nabla^2 \bm{u}^{n-1/2} - \nabla{p}^{n-3/2} + \bm{f}, \quad \bm{u}^*=0 \text{ for } x=\pm1   \label{eq:NSCN} \\
ii)&\,\,\begin{cases}
  \frac{\bm{u}^{n}-\bm{u}^*}{\triangle t} &= -\nabla p^* \\
  \nabla \cdot \bm{u}^{n} &= 0, \quad \bm{u}^n\cdot \bm{n} = 0,\text{ for } x=\pm1
  \end{cases} \label{eq:momupdate}\\
  iii)&\,\,p^{n-1/2}=p^{n-3/2}+p^* \label{eq:pupdate} 
\end{align}
Here $\bm{u}^{n-1/2} = 0.5(\bm{u}^*+\bm{u}^{n-1})$, $\bm{u}^*$ is a tentative velocity, $p^*$ is a pressure correction, and the explicit convection is computed as $\bm{N}^{n-1/2}=1.5(\bm{u}^{n-1}\cdot \nabla) \bm{u}^{n-1} - 0.5(\bm{u}^{n-2}\cdot \nabla) \bm{u}^{n-2}$. The second step is computed by solving a Poisson equation for the pressure correction
\begin{equation}
  \nabla^2p^* = \frac{\nabla \cdot \bm{u}^*}{\triangle t}
\end{equation}
and subsequently updating the velocity through (\ref{eq:momupdate}).

All three velocity components use the Dirichlet basis (\ref{eq:u_sol}), whereas the pressure is searched for in $\N{V}_N$
\begin{equation}
p(\bm{x}, t) = \frac{1}{N_yN_z}\sum_{\bm{k} \in \N{K}_N} \hat{p}_{\bm{k}}(t) \N{\psi}_{\bm{k}}(\bm{x}), \label{eq:p_sol}
\end{equation}

The spectral Galerkin method becomes:

1) Find $\bm{u}^*$ in $\D{V}_N^3$ by solving
\begin{multline}
 \frac{1}{\triangle t}\left<\bm{u}^*-\bm{u}^{n-1}, \bm{\D{\psi}}\right>_w + \left<\bm{N}^{n-1/2}, \bm{\D{\psi}} \right>_w   = \nu \left<\nabla^2 \bm{u}^{n-1/2}, \bm{\D{\psi}} \right>_w \\ - \left<\nabla{p}^{n-3/2}, \bm{\D{\psi}} \right>_w + \left<\bm{f}, \bm{\D{\psi}} \right>_w \,\,  \forall \, \bm{\D{\psi}} \in \D{V}_N^3. \label{eq:tentative}
\end{multline}

2) Compute the pressure correction $p^*\in \N{V}_N$ by solving
\begin{equation}
\left< \nabla^2 p^*, \N{\psi} \right>_w = \frac{1}{\triangle t} \left<\nabla \cdot \bm{u}^*, \N{\psi} \right>_w\,\, \forall \, \N{\psi} \in \N{V}_N. \label{eq:pcorr}
\end{equation}

>>>>

\section{Preliminaries}
The Navier-Stokes equations used to describe turbulent flow in a doubly 
periodic channel can be written as
\begin{align}
 \frac{\partial \bm{u}}{\partial t}   &= \bm{\mathcal{H}} + \nu 
 \nabla^2 \bm{u} - \nabla{p}, \notag \\
 \nabla \cdot \bm{u} &= 0, \label{eq:NS}
\end{align}
where $\bm{u}(\bm{x}, t)=(u, v, w)$ is the velocity vector, $\bm{x}=(x, y, z)$ 
and $t$ are position and time, $\bm{\mathcal{H}}(\bm{x}, t)$ represents the 
nonlinear convection and the mean pressure gradient (the driving force), $\nu$ 
is a constant dynamic viscosity and $p$ is the instantaneous pressure 
fluctuation normalized by a constant density. The 
computational domain is $\Omega=[-1, 1]\times [0, L_y] \times [0, L_z]$ with 
channel walls located at $x=\pm 1$ such that no-slip applies at $ \bm{u}(\pm 1, 
y, z, t) = 0$. The walls are spanning the $y-z$ plane and the equations are 
periodic in the $y$ and $z$ directions.

The Navier Stokes equations are discretized using Fourier basis functions for the periodic directions, and a combination of Chebyshev polynomials in the wall normal direction. Three different sets of basis functions and function spaces are relevant for the wall normal direction
\begin{align}
&  \phi_k(x) = T_k(x), & V_{N_x} &= \text{span}\{\phi_0, \phi_1,\ldots, \phi_{N_x}\} \label{eq:Tk}\\
& \D{\phi}_k(x) = T_k(x) - T_{k+2}(x), & \D{V}_{N_x} &= \text{span} \{ \D{\phi}_0, \D{\phi}_1, \ldots, \D{\phi}_{N_x-2} \} \label{eq:phiD}\\
& \N{\phi}_k(x) = T_k(x) - \frac{2(k+2)}{k+3} T_{k+2}(x) + 
\frac{k+1}{k+3} T_{k+4}, & \N{V}_{N_x} &= \text{span} \{\N{\phi}_0, \N{\phi}_1, 
\ldots, \N{\phi}_{N_x-4} \} \label{eq:phiN} 
\end{align}
where $T_k(x)$ is the $k$'th degree Chebyshev polynomial of the first kind. The 
basis functions and function spaces in (\ref{eq:phiD}) and (\ref{eq:phiN}) were 
suggested by Shen, and satisfy, respectively, the boundary conditions 
$\D{\phi}_k(\pm 1) = 0$, $\N{\phi}_k(\pm 1)=0$ and $\N{\phi}'_k(\pm 1)=0$. The 
last two function spaces may alternatively be written as $\D{V}_{N_x} = \{v \in 
V_{N_x}: v(\pm 1)=0 \}$ and $\N{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1) = v'(\pm 
1) = 0 \}$.

Three-dimensional basis functions and function spaces that are periodic in $y$ 
and $z$ directions, can now be defined as
\begin{align}
  \psi_{\bm{k}}(\bm{x}) = \phi_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad V_N &= \text{span} \{ \psi_{\bm{k}}(\bm{x}):\, \bm{k} \in K_N  \}, \\
  \D{\psi}_{\bm{k}}(\bm{x}) = \D{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \D{V}_N &= \text{span} \{ \D{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \D{K}_N  \}, \\
  \N{\psi}_{\bm{k}}(\bm{x}) = \N{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \N{V}_N &= \text{span} \{ \N{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \N{K}_N  \},
\end{align}
where $\imath=\sqrt{-1}$ and 
\begin{align}
K_N = \Big\{\bm{k} \in \mathcal{Z}\times\mathcal{R}^2 / &(l, \underline{m}, \underline{n}) \in \left( l, \frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
 &(l, m, n) \in  [0, \ldots, N_x] \times [-\frac{N_y}{2},\ldots, \frac{N_y}{2}-1] \times[-\frac{N_z}{2},\ldots, \frac{N_z}{2}-1] \Big\}. \label{eq:wavenumbermesh}
\end{align}
$\D{K}_N$ and $\N{K}_N$ are both equal to $K_N$ except from index $l$ that 
ranges from $[0, 1, \ldots, N_x-2]$ and $[0, 1, \ldots, N_x-4]$ for $\D{K}_N$ 
and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). The 
computational mesh has in real space been created using $N = (N_x, N_y, N_z)$ 
intervals, where the two periodic directions use uniform intervals. The 
computational mesh is given as
\begin{align}
  X_N = \Big\{ \bm{x} \in \mathcal{R}^3/&(x_i, y_j, z_k) \in \left( h(i), \frac{jL_y}{N_y}, \frac{kL_z}{N_z} \right), \text{where} \notag \\
  &(i, j, k) \in [0, 1, \ldots, N_x] \times [0, 1, \ldots, N_y-1] \times [0, 1, \ldots, N_z-1] \Big\} \label{eq:Xn}
\end{align}
where $x_i = h(i)$ represents either
\begin{equation}
 h(i) = \begin{cases}
   \cos \left(\frac{i \pi }{N_x} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Lobatto points}, \\
   \cos \left(\frac{(2i +1)\pi}{2N_x+2} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Chebyshev points}. \\
 \end{cases}
\end{equation}

For the Navier Stokes equations we look for solutions of the velocity 
components of the form
\begin{align}
u(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \N{K}_N} \hat{u}_{\bm{k}}(t) 
\N{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_solx} \\
v(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \D{K}_N} \hat{v}_{\bm{k}}(t) 
\D{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_soly} \\
w(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \D{K}_N} \hat{w}_{\bm{k}}(t) 
\D{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_solz}
\end{align}
where $\hat{u}_{\bm{k}}(t) = \hat{u}(l, {m}, {n}, t)$ are the expansion 
coefficients for the velocity component in $x$-direction (and similar for the 
other two components) and the scaling by $N_y$ and $N_z$ is merely for 
convenience 
and compliance with the definition used for the inverse discrete Fourier 
transform. Note that from now on we will simply use the notation $\hat{u}$ for 
$\hat{u}_{\bm{k}}(t)$, when it is possible to simplify without loss of clarity. 
Likewise we will simply use $u$ for $u(\bm{x}, t)$.

If the expansion coefficients $\hat{u}$  are known for the entire wavenumber 
mesh (\ref{eq:wavenumbermesh}), then the expression (\ref{eq:u_solx}) may be 
evaluated on the mesh (\ref{eq:Xn}) using fast Fourier and fast Shen inverse 
transforms as
\begin{align}
u(x_i, y_j, z_k, t) &= \underbrace{\frac{1}{N_z}\sum_n 
\underbrace{\frac{1}{N_y} \sum_m \underbrace{\sum_l 
\hat{u}(l,m,n)\N{\phi}_l(x_i)}_{\N{\mathcal{S}}_x^{-1}} e^{\imath \underline{m} 
y_j}}_{\mathcal{F}_y^{-1}} e^{\imath \underline{n} z_k}}_{\mathcal{F}_z^{-1}} , 
\quad \forall \, \bm{x} \in X_N, \notag \\
  u &= \N{\mathcal{T}}^{-1}(\hat{u}) =  \mathcal{F}^{-1}_z (\mathcal{F}^{-1}_y 
  (\N{\mathcal{S}}^{-1}_x (\hat{u}))), \quad \forall \, \bm{x} \in X_N.  
  \label{eq:ifft} 
\end{align}
Here $\N{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse 
transform, $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represents inverse 
Fourier transforms along directions $y$ and $z$ respectively, and the inverse 
Shen transform $\N{\mathcal{S}}_{x}^{-1}$ is performed along the wall normal 
$x$ direction. Note that the transforms are applied sequentially to the entire 
computational mesh. The inverse Shen transform may be computed using fast 
Chebyshev transforms for all the three bases in (\ref{eq:Tk}, \ref{eq:phiD}, 
\ref{eq:phiN}). The transforms are slightly different for the three bases and 
${\mathcal{S}}, \D{\mathcal{S}}$ and $\N{\mathcal{S}}$ are used to distinguish 
between them with obvious notation. Python implementations of forward and 
inverse Shen transforms and scalar products are given in the Appendix.

A three-dimensional scalar product in the weighted $[L^2_w(\Omega)]^3$ space is defined as
\begin{align}
 \left<u, v\right>_w &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}w\,dxdydz, \notag \\
   &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}\,dx_wdydz,
\end{align}
where $v^*$ is the comlex conjugate of $v$ and the weights $w$ are unity for periodic directions and  $w=1/\sqrt{1-x^2}$ for the inhomogeneous direction, and we have used $dx_w = dx/\sqrt{1-x^2}$. In this work we will make use of the discrete weighted $[l^2_w(\Omega)]^3$ space, where quadrature is employed for integration. As such the scalar product corresponds to sums that may be computed using fast transforms
\begin{align}
 \left< u, \N{\psi}_{\bm{k}} \right>_w &= h \underbrace{\sum_i 
 \underbrace{\sum_j \underbrace{\sum_k u(x_i, y_j, z_k, t)  e^{-\imath 
 \underline{n} z_k}}_{\mathcal{F}_n}  e^{-\imath \underline{m} y_j} 
 }_{\mathcal{F}_m} \N{\phi}_{l}(x_i) w_i}_{\N{\mathcal{S}}_l},   \notag \\
  &=  h \N{\mathcal{S}}(u) = h \N{\mathcal{S}}_{l} (\mathcal{F}_{{m}} 
  (\mathcal{F}_{{n}}(u))), \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:sst1}
\end{align}
Here $h\N{\mathcal{S}}$ denotes the complete three dimensional scalar product 
and $h = L_yL_zN_y^{-1}N_z^{-1}$ is a constant. $\mathcal{F}_{{n}}$ and 
$\mathcal{F}_{{m}}$ represent discrete Fourier transforms in $z$- and 
$y$-directions, respectively, and $\N{\mathcal{S}}_{l}(\cdot) = (\cdot, 
\N{\phi}_l)_w$ is the Shen scalar product in the $x$-direction for the 
$\N{V}_N$ space. The weights, $w_i$, required for the Shen transforms are 
defined as
\begin{equation}
 w_i = \begin{cases}
       \frac{\pi}{c_i N_x} &\forall \, i=0,1,\ldots, N_x \quad  \text{for 
       Gauss-Lobatto points},\\
       \frac{\pi}{N_x+1} &\forall \, i=0,1,\ldots, N_x  \quad \text{for 
       Gauss-Chebyshev points},      
 \end{cases}
\end{equation}
where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$.

The scalar product  $h\N{\mathcal{S}}$ in (\ref{eq:sst1}) does not represent a 
complete transform. To find a transformation between physical $u$ and spectral 
$\hat{u}$ we make use of Eq. (\ref{eq:u_solx}) directly and use the discrete 
orthogonality of the Fourier basis functions
\begin{align}
\left<u, \N{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< 
\sum_{\bm{m}=(q,r,s) \in \N{K}_N} \hat{u}_{\bm{m}} \N{\psi}_{\bm{m}}, 
\N{\psi}_{\bm{k}} \right>_w, \notag \\
           &= h \sum_{q=0}^{N_x-4} \sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) 
           \N{\phi}_{l}(x_i) w_i \, \hat{u}(q, {m}, {n}, t), \notag \\
           &= h \sum_{p=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t), 
           \label{eq:sst2}
\end{align}
where the scalar product $\N{B}_{lq} = (\N{\phi}_q, \N{\phi}_l)_w = 
\sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) \N{\phi}_{l}(x_i) w_i$ is a 
pentadiagonal mass matrix. Likewise, for the remaining two velocity components, 
the corresponding mass matrix, $\D{B}_{lq} = (\D{\phi}_q, \D{\phi}_l)_w$, is 
tridiagonal. The complete transformation is now obtained by setting Eq. 
(\ref{eq:sst2}) equal to (\ref{eq:sst1}) and solving for $\hat{u}$. Moving to 
matrix notation we get
\begin{align}
h\sum_{q=0}^{N_x-4} \N{B}_{lq}\, \hat{u}(q, {m}, {n}, t) &= 
h\N{\mathcal{S}}(u)(l, m, n, t), \notag \\
 \hat{u}_{\bm{k}}(t) &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u), 
 \quad \forall \, \bm{k} \in \N{K}_N.
\end{align}
where $\N{\mathcal{T}}(u)$ denotes the complete transformation, such that $u = 
\N{\mathcal{T}}^{-1}(\N{\mathcal{T}}(u))$. Note that since $\N{B}$ assembles to 
a pentadiagonal matrix the solution ($\N{B}^{-1}$) can be obtained very fast 
and 
the complete 
transformation thus requires a fast Chebyshev transform ($O(N_x \log N_x)$) and 
a fast linear algebraic solve ($O(N_x)$). Similar transforms $\mathcal{T}$ 
and $\D{\mathcal{T}}$ are defined for the two other bases (\ref{eq:Tk}) and 
(\ref{eq:phiD}), using mass matrices $B_{lq}=(T_q, T_l)_w$ (diagonal) and 
$\D{B}_{lq}=(\D{\phi}_l, \D{\phi}_q)_w$ (tridiagonal) and scalar products 
$\mathcal{S}_l(\cdot) = (\cdot, T_l)_w$ and $\D{\mathcal{S}}_l(\cdot) = (\cdot, 
\D{\phi}_l)_w$.
